---
title: "week-8-hw-eeb177"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Part 1
## 6. Canids

```{r}
library(ggplot2)
setwd("/home/eeb177-student/Desktop/eeb-177/homework")

canids <- read.csv("/home/eeb177-student/Desktop/eeb-177/homework/canid-ranges.csv", header =F, as.is = T)

names(canids) <- c("genus", "species", "minage", "maxage")
head(canids)

#this gives us the data in a nice form for R
```

```{r}
canid_occ <- ggplot(canids, aes( species, ymin = maxage, ymax=minage, colour = genus))
canid_occ <- canid_occ + geom_linerange()
canid_occ <- canid_occ + theme(legend.position="none") #gets rid of legend
canid_occ <- canid_occ + coord_flip() #flips the coordinats horizontally
canid_occ <- canid_occ +  theme(axis.text.y = element_text(size=3)) # resizes text
canid_occ <- canid_occ + theme(axis.ticks.y=element_blank()) # gets rid of tick marks on y axis
canid_occ <- canid_occ + scale_y_continuous(limits=c(0, 40), expand = c(0, 0), breaks=c(0, 10, 20, 30, 40)) # gets rid of space on right and left of the graph to make the graph itself larger
canid_occ <- canid_occ + labs(title = "Canid Fossil Occurrences", x = "Species", y = "Ma ago") + theme(plot.title = element_text(hjust = 0.5, size=22, face = "bold"), axis.title =element_text(size=20))  # adds titles and labels to the graph
canid_occ
```

``` {r}
ggsave(filename = "canid-occ.pdf", plot = canid_occ)
```
## 7.
  1.  What taxonimic group is displayed?
```{r}
print("Canids")
```
  2.  How many unique fossil speies are contained within it?
```{r}
cat("Unique species:", nrow(canids))
# 172
```
  3.  How many unique fossil genera?
```{r}
list_of_genera <- unique(canids$genus)
cat("Number of Genera:", length(list_of_genera))
# 58
```
  4.  What species has the longest duration?
```{r}
canids$duration <- canids$maxage - canids$minage
#gives the data frame a new row with the duration
max_duration_taxa<- which.max(canids$duration) 
#this gives the row of the max value
cat(canids[max_duration_taxa,2], canids[max_duration_taxa,5], "Ma")

#use this to call the appropriate taxa within the data frame
```
# Nautilaceae
```{r}

nautilaceae <- read.csv("/home/eeb177-student/Desktop/eeb-177/homework/nautilaceae-range.csv", header =F, as.is = T)

names(nautilaceae) <- c("genus", "species", "minage", "maxage")
head(nautilaceae)
```

```{r}
nautilaceae_occ <- ggplot(nautilaceae, aes( species, ymin = maxage, ymax=minage, colour = genus))
nautilaceae_occ <- nautilaceae_occ + geom_linerange()
nautilaceae_occ <- nautilaceae_occ + theme(legend.position="none") #gets rid of legend
nautilaceae_occ <- nautilaceae_occ + coord_flip() #flips the coordinats horizontally
nautilaceae_occ <- nautilaceae_occ +  theme(axis.text.y = element_text(size=1)) # resizes text
nautilaceae_occ <- nautilaceae_occ + theme(axis.ticks.y=element_blank()) # gets rid of tick marks on y axis
 nautilaceae_occ <- nautilaceae_occ + scale_y_continuous(limits=c(0, 250), expand = c(0, 0), breaks=c(0, 50, 100, 150, 200, 250)) # gets rid of space on right and left of the graph to make the graph itself larger
nautilaceae_occ <- nautilaceae_occ + labs(title = "Nautilaceae Fossil Occurrences", x = "Species", y = "Ma ago") + theme(plot.title = element_text(hjust = 0.5, size=22, face = "bold"), axis.title =element_text(size=20))  # adds titles and labels to the graph
nautilaceae_occ
```

``` {r}
ggsave(filename = "nautilaceae_occ.pdf", plot = nautilaceae_occ)
```

## 7.
  1.  What taxonimic group is displayed?
```{r}
print("Nautilaceae")
```
  2.  How many unique fossil speies are contained within it?
```{r}
cat("Unique Species:", nrow(nautilaceae))
# 471
```
  3.  How many unique fossil genera?
```{r}
list_of_genera <- unique(nautilaceae$genus)
cat("Number of Genera:", length(list_of_genera))
# 39
```
  4.  What species has the longest duration
```{r}
nautilaceae$duration <- nautilaceae$maxage - nautilaceae$minage
#gives the data frame a new row with the duration
max_duration_taxa<- which.max(nautilaceae$duration) 
#this gives the row of the max value
cat(nautilaceae[max_duration_taxa,2], nautilaceae[max_duration_taxa,5], "Ma")
#use this to call the appropriate taxa within the data frame
```


# Part 2

## 1.
```{r}
cats <- read.csv(file = "feline-data.csv")
age <- c(4,5,8)
cats <- cbind(cats, age)
```

```{r}
newRow <- list("tortoiseshell", 3.3, TRUE, 9)
cats <- rbind(cats, newRow)
```

```{r}
levels(cats$coat)


levels(cats$coat) <- c(levels(cats$coat), 'tortoiseshell')
cats <- rbind(cats, list("tortoiseshell", 3.3, TRUE, 9))

cats
```

```{r}
str(cats)
```

```{r}
cats$coat <- as.character(cats$coat)
str(cats)
```

```{r}
# remove the 4th column
cats[-4,]
```

```{r}
cats <- na.omit(cats)
```

```{r}
cats <- rbind(cats, cats)
cats
rownames(cats) <- NULL
cats
```

### Challenge 1
```{r}
df <- data.frame(first_name = c("Hayden"), last_name = c("Speck"), lucky_number = c(657890), stringsAsFactors = F)
df <- rbind(df, list("Jean-Erique", "Namme", 0))
df <- cbind(df, coffee = c(FALSE, TRUE))
df
```


```{r}
gapminder <- read.csv("gapminder-FiveYearData.csv")
```

```{r}
str(gapminder)
```
```{r}
typeof(gapminder$country)
```
```{r}
str(gapminder$country)
```
```{r}
length(gapminder)
```
```{r}
typeof(gapminder)
```
```{r}
nrow(gapminder)
```
```{r}
ncol(gapminder)
```
```{r}
dim(gapminder)
```
```{r}
colnames(gapminder)
```
```{r}
head(gapminder)
```
### Challenge 2
```{r}
source(file = "~/scripts/load-gapminder.R")
```

### Challenge 3
gapminder is data frame
1704 rows and 6 columns
Country and Continent are factors
year is vector (integer)
pop, lifeExp and gdp are vectors (numeric)



## 2. Subsetting Data
```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
x
```

```{r}
print(x[1])
print(x[4])
print(x[c(1,3)])
x[1:4]
```

```{r}
print(x[c(1,1,3)])
print(x[6])
x[0]
```

```{r}
print(x[-2])
print(x[c(-1,-5)])
```

```{r}
x<-x[-4]
x
```
### Challenge 1
```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)

print(x[2:4])
print(x[c(-1,-5)])
print(x[c(2,3,4)])
```

```{r}
x[c("a", "c")]
```

```{r}
x[-which(names(x) == "a")]
```
### Challenge 3
```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)
test_type <- x[-which(names(x) == "g")]
#expect a null result
#instead got a named numeric(0)
class(test_type)
#returns a numeric
```
A) True
B) False
C) true
D)false
E)false

```{r}
x <- 1:3
names(x) <- c("a", "a", "a")
x
```

```{r}
```

```{r}
names(x) == c('a', 'c')
```



```{r}
print(x[c(TRUE, TRUE, FALSE, FALSE)])
print(x[c(TRUE, FALSE)])
```

```{r}
x[x>7]
```
### Challenge 3
```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')

x_subset <- x[x<7 & 4<x]
print(x_subset)
```


```{r}
f <- factor(c("a", "a", "b", "c", "c", "d"))
f[f == "a"]
```
```{r}
f[f %in% c("b", "c")]
```
```{r}
f[1:3]
```
```{r}
set.seed(1)
m <- matrix(rnorm(6*4), ncol=4, nrow=6)
m[3:4, c(3,1)]

```
```{r}
m[, c(3,4)]
```
```{r}
m[3, , drop=FALSE]
```
```{r}
m[, c(3,6)]
```
```{r}
m[5]
```
```{r}
matrix(1:6, nrow=2, ncol=3)
```
```{r}
matrix(1:6, nrow=2, ncol=3, byrow=TRUE)
```

### Challenge 4
```{r}
m <- matrix(1:18, nrow=3, ncol=6)
print(m)
```

```{r}
m[2,4,2,5]

m[2:5]

m[4:5,2]

m[2,c(4,5)]
```
A: No, too many dimensions
B: No will grab a range of values
C: No subscript out of bounds
D: Yes, will grab 2,4 then 2,5

```{r}
xlist <- list(a = "Software Carpentry", b = 1:10, data = head(iris))
xlist[1]
```
```{r}
xlist[1:2]
```
```{r}
xlist[[1]]
```
```{r}
xlist[[1:2]]
```
```{r}
xlist[[-1]]
```
```{r}
xlist[["a"]]
```
```{r}
xlist$data
```
### Challenge 5
```{r}
xlist <- list(a = "Software Carpentry", b = 1:10, data = head(iris))

xlist$b[2]

```

### Challenge 6
```{r}
mod <- aov(pop ~ lifeExp, data=gapminder)

attributes(mod)
mod$df.residual
```

```{r}
head(gapminder[3])
```
```{r}
head(gapminder[["lifeExp"]])
```
```{r}
head(gapminder$year)
```
```{r}
gapminder[1:3,]
```
```{r}
gapminder[3,]
```
### Challenge 7
```{r}
# gapminder[gapminder$year = 1957,]
gapminder[gapminder$year == 1957,]
```
```{r}
#gapminder[,-1:4]
gapminder[,-c(1:4)]
```
```{r}
#gapminder[gapminder$lifeExp > 80]
gapminder[gapminder$lifeExp > 80,]
```
```{r}
#gapminder[1, 4, 5]
gapminder[1, c(4, 5)]
```
```{r}
 # gapminder[gapminder$year == 2002 | 2007,]
gapminder[gapminder$year == 2002 | gapminder$year == 2007,]
```

### Challenge 8
1. gapminder[1:20, ] the columns and rows of the frame
gapminder[1:20] only gives one dimension of the frame
```{r}
gapminder[1:20]
```
```{r}
gapminder_small <- gapminder[c(1:9, 19:23),]
gapminder_small
```
```{r}
```
```{r}
```
```{r}
```
```{r}
```
```{r}
```
```{r}
```
```{r}
```
```{r}
```
```{r}
```
```{r}
```
